---
title: "p8105_hw2_jl6647"
author: "Jiatong LI"
date: "2023-09-27"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
```
## HW2 Problem 2

### Read and clean Mr. Trash Wheel's data

* specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
* use reasonable variable names 
* omit rows that do not include dumpster-specific data

### Mr. Trash Wheel
```{r}
trash_df = read_excel(
  "202207 Trash Wheel Collection Data.xlsx",
  range = "A2:N549",
  sheet = "Mr. Trash Wheel"
) |> 
  janitor::clean_names() |>
  mutate(year = as.integer(year)) |> 
  mutate(homes_powered = weight_tons*500/30) |> 
  mutate(wheel_type = "mr") |> 
  drop_na(dumpster)

```
### Professor Trash Wheel
```{r}
professor_df = read_excel(
  "202207 Trash Wheel Collection Data.xlsx",
  sheet = "Professor Trash Wheel",
  range = "A2:M96"
) |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(homes_powered = weight_tons*500/30) |> 
  mutate(wheel_type = "professor")
```
### Gwynnda Trash Wheel
```{r}
gwynnda_df = read_excel(
  "202207 Trash Wheel Collection Data.xlsx",
  sheet = "Gwynnda Trash Wheel",
  range = "A2:K109"
) |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(homes_powered = weight_tons*500/30) |> 
  mutate(wheel_type = "gwynnda")
```

### Combine datasets together

```{r}
combined_dataset = bind_rows(trash_df,professor_df,gwynnda_df) |> 
                   relocate(wheel_type)
print(combined_dataset)
```

### Describe the data

The number of observations in the combined dataset: `r nrow(combined_dataset)`

The number of variables in the combined dataset: `r ncol(combined_dataset)`

Names of key variables: `r names(combined_dataset)`

Total weight of trash collected by Professor Trash Wheel: `r sum(professor_df$weight_tons)`

Total number of cigarette butts collected by Gwynnda in July of 2021: `r sum(gwynnda_df$cigarette_butts[which(gwynnda_df$month == "July" & gwynnda_df$year == 2021)])`






